Insights Into Computational Reproducibility in Plant Pathology and a Way Forward

### 2021-06-14 analysis/paper/paper.Rmd
Each sentence on it's own line
evaluate library using standard evaluation
Replace tables with four-panel figure

library(Reproducibility.in.Plant.Pathology)
```

@@

Most articles did not make any computational methods available in any fashion with two, 0.007%, classing as "3", which was the highest score available.
Two hundred eighty-seven articles, 96%, were classed as "0" and nine articles, 0.03%, appeared to not use any computational methods.

Likewise, data availability was mostly not available with 243 or 81.5% scoring "0" where the data was not available or mentioned in the article.
However, more articles, 29 or 9.7%, scored "3" than "1" or "2" combined, with two articles not producing shareable data (Table \@ref(tab:da)).

More articles did a good job of using software that was reasonably available to anyone with 206 scoring "2" or "3" (Table \@ref(tab:sa)).

Software citations were also reasonably well reported in the literature with 120 scoring "2" where the main software was cited with version numbers and 57 scoring "3" where full citations including R packages or the SAS PROC used were cited where appropriate (Table \@ref(tab:sc)).

@@

```{r scoring, tab.id="scoring", tab.cap ="Articles were scored in four areas for reproducibility on a scale of 0 to 3 by four independent assessors.", echo=FALSE, message=FALSE, warning=FALSE}
scoring_criteria %>% 
  rename(`Index Value` = Index.Value) %>% 
  flextable() %>% 
  colformat_num(na_str = "NA") %>% 
  set_table_properties(layout = "autofit")
```

@@

```{r software_avail, tab.id="sa", tab.cap="Article ratings for software availability where '0', not available or not mentioned in publication to '3', preferably open source and freely available.", echo=FALSE, message=FALSE, warning=FALSE}
import_notes() %>%
  tabyl(software_avail) %>%
  adorn_pct_formatting() %>% 
  select(-valid_percent) %>% 
  rename(`Software Availability` = software_avail,
         Percent = percent) %>% 
  flextable() %>%
  colformat_num(na_str = "NA") %>% 
  autofit()
```

```{r software_cite, tab.id="sc", tab.cap="Article ratings for software citations where '0', not specified or mentioned in publication to '3', fully cited with proper version numbers and all procedures detailed.", echo=FALSE, message=FALSE, warning=FALSE}
import_notes() %>%
  tabyl(software_cite) %>%
  adorn_pct_formatting() %>% 
  select(-valid_percent) %>% 
  rename(`Software citation` = software_cite,
         Percent = percent) %>% 
  flextable() %>%
  colformat_num(na_str = "NA") %>% 
  autofit()
```

```{r data_available, tab.id="da", tab.cap="Article ratings for data availability where '0', not specified or mentioned in publication to '3', Freely available online to anonymous users for foreseeable future (_e.g._, archived using Zenodo, dataverse or university library or some other proper archiving system).", echo=FALSE, message=FALSE, warning=FALSE}
import_notes() %>%
  tabyl(data_avail) %>%
  adorn_pct_formatting() %>% 
  select(-valid_percent) %>% 
  rename(`Data availability` = data_avail,
         Percent = percent) %>% 
  flextable() %>%
  colformat_num(na_str = "NA") %>% 
  autofit()
```

```{r software_used, tab.id="su", tab.cap="Tally of reproducibility scores where 0 is not reproducible and 12 is fully reproducible, _e.g._, all software cited, software used are freely available to anyone, computational methods (statistical analysis or other computational code) are freely available for download and use, and raw data are available for download and use.", echo=FALSE, message=FALSE, warning=FALSE}
library("tidyr")

@@

```{r reproducibility_score, tab.id="reproducibility", tab.cap="Tally of reproducibility scores where 0 is not reproducible and 12 is fully reproducible, _e.g._, all software cited, software used are freely available to anyone, computational methods (statistical analysis or other computational code) are freely available for download and use, and raw data are available for download and use.", echo=FALSE, message=FALSE, warning=FALSE}
import_notes() %>%
  tabyl(reproducibility_score) %>%
  adorn_pct_formatting(digits = 1) %>%
  rename(Score = reproducibility_score, Percent = percent) %>%
  flextable() %>%
  autofit()
```

@@

```{r workflow-dia, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='An open and reproducible research workflow.', cache=FALSE, fig.env='figure'}
library("DiagrammeR")
library("DiagrammeRsvg")
library("rsvg")

@@

```{r article_scores, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Aggregated article scores for each of the four categories. In subfigure `a`, `NA` indicates that no computational methods were used to generate the data, _e.g._ only PCR results were reported with no statistical analysis performed.', cache=FALSE, fig.env='figure'}
library("ggplot2")
library("patchwork")
rrpp <- import_notes()
theme_set(theme_classic())
a <- ggplot(rrpp, aes(y = as.factor(comp_mthds_avail))) +
  geom_bar() +
  ggtitle("Computational Methods Availability") +
  ylab("Score") +
  xlab(NULL)

>>>>

library("Reproducibility.in.Plant.Pathology")
```

@@

Most articles did not make any computational methods available in any fashion with two, 0.007%, classing as "3", which was the highest score available (Figure \@ref(fig:article-scores)a).
Two hundred eighty-seven articles, 96%, were classed as "0" and nine articles, 0.03%, appeared to not use any computational methods.

More articles did a good job of using software that was reasonably available to anyone with 206 scoring "2" or "3" (Figure  \@ref(fig:article-scores)b).

Software citations were also reasonably well reported in the literature with 120 scoring "2" where the main software was cited with version numbers and 57 scoring "3" where full citations including R packages or the SAS PROC used were cited where appropriate (Figure \@ref(fig:article-scores)c).

However, data availability was mostly not available with 243 or 81.5% scoring "0" where the data was not available or mentioned in the article.
However, more articles, 29 or 9.7%, scored "3" than "1" or "2" combined, with two articles not producing shareable data (Figure  \@ref(fig:article-scores)d).

@@

```{r scoring, tab.id="scoring", tab.cap ="Articles were scored in four areas for reproducibility on a scale of '0' to '3' by four independent assessors to rank articles in each of four areas, 'Computational methods availability', 'Data availability', 'Software availability' and 'Software citation'.", echo=FALSE, message=FALSE, warning=FALSE}
scoring_criteria %>% 
  rename(`Index Value` = Index.Value) %>% 
  flextable() %>% 
  colformat_num(na_str = "NA") %>% 
  set_table_properties(layout = "autofit")
```

@@

```{r software_used, tab.id="su", tab.cap="The top ten sofware programs that were found to be used in the papers that were surveyed.", echo=FALSE, message=FALSE, warning=FALSE}
library("tidyr")

@@

```{r reproducibility_score, tab.id="reproducibility", tab.cap="Tally of reproducibility scores where '0' is not reproducible and '12' is fully reproducible, _e.g._, all software cited, software used are freely available to anyone, computational methods (statistical analysis or other computational code) are freely available for download and use, and raw data are available for download and use.", echo=FALSE, message=FALSE, warning=FALSE}
import_notes() %>%
  tabyl(reproducibility_score) %>%
  adorn_pct_formatting(digits = 1) %>%
  rename(Score = reproducibility_score, Percent = percent) %>%
  mutate(Score = round(Score, digits = 1)) %>% 
  flextable() %>%
  autofit()
```

@@

<!-- @Emerson, can you please have a look at this caption, if I'm not mistaken, this is your diagram design. Can you confirm that what I've written is accurate and describes the workflow? -->

```{r workflow-dia, echo=FALSE, message=FALSE, warning=FALSE, fig.cap='An example open and reproducible research workflow. Starting with the question, determine the methodology, describe it, make it available and cite it. Data are used in analysis and any binary files or code are made available as supplements to the manuscript. Source code and raw data are made available in a public repository, preferably with version control for tracking changes through time and a DOI for final released products.', cache=FALSE, fig.env='figure'}
library("DiagrammeR")
library("DiagrammeRsvg")
library("rsvg")

@@

```{r article-scores, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Aggregated article scores for each of the four categories, a), computational methods availability 0 - not available or not mentioned to 3 - freely available online to anonymous users. b) Shows 'Software availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Expensive proprietary software that only institutions would typically purchase'; '2' was 'Used proprietary software that most individuals can afford (e.g., Excel) or SAS that is now available free for use with a university login'; and '3' was 'Used entirely open source and free software (e.g., R, Julia, Python)'; 'NA' indicates that no specialised software was necessary or used. c) Displays 'Software citated', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Software mentioned by name only'; '2' was 'Software cited with version number'; and '3' was 'All software components (SAS PROCs, R, Julia or Python packages, etc.) cited'; 'NA' indicates that no specialised software was necessary or used. d) Shows 'Data availability', where '0' was 'Not available or not mentioned in the publication'; '1' was 'Available upon request to the author; '2' was 'Online, but inconvenient or non-permanent (e.g., login needed, paywall, FTP server, personal lab website that may disappear, or may have already disappeared)'; and '3' was 'Freely available online to anonymous users for foreseeable future (e.g., archived using Zenodo, dataverse or university library or some other proper archiving system)'; 'NA' indicates that no data were generated, _e.g._ a methods paper.", cache=FALSE, fig.env='figure'}
library("ggplot2")
library("patchwork")
rrpp <- import_notes()
theme_set(theme_classic())
a <- ggplot(rrpp, aes(y = as.factor(comp_mthds_avail))) +
  geom_bar() +
  ggtitle("Computational Methods Availability") +
  ylab("Score") +
  xlab(NULL)
